% Briefly describe your existing extramural support and plans for generating follow-on funding.


\input{preamble.tex}
% even page, right position (inner)
\rehead[]{Dav Clark -- \textbf{Funding} -- Data Science for Social Impact}
% odd page, left position (inner)
\lohead[]{Dav Clark -- \textbf{Funding} -- Data Science for Social Impact}

\begin{document}

I am currently employed by the D-Lab and paid out of central campus funds. My
current appointment is for one year and started in February, 2014. I cannot
guarantee that this position will be extended, but the general opinion in the
D-Lab is that it's likely. As mentioned in Cathryn Carson's excerpt, she is
unable to guarantee this however, as she wishes to allow the incoming faculty
director, Justin McCrary, to have a say in how the D-Lab is run (and in
particular how it's staffed).

In the unlikely event that I am \emph{not} renewed at the D-Lab, there are a
number of potential units on campus that I might continue working with. I
imagine that having a pre-existing BIDS fellowship would further increase my
ability to obtain such a position. For example, I have already spoken with
Patrick Schmitz (associate director of Shared Research Computing) about working
with his group on cross-pollinating advanced computational resources across
campus in the form of consulting as well as physical infrastructure. Moreover, I
expect to have the support of Cathryn Carson (among others) in retaining a
position at UC Berkeley.

\end{document}
