% (1 page maximum) Please provide an overview of the funding that will be used to
% cover the 50% portion of the salary (incl. benefits) that will not be provided
% by BIDS. The 50% funding provided by BIDS will be administered in a flexible
% fashion and averaged over the entire two-year period (i.e. could translate into
% one year of salary provided by BIDS and one by the matching source, or a 50/50
% split arrangement over the course of the two years.) If you are a faculty member
% seeking a position as a Data Science Fellow please describe how the requested
% research support, teaching relief and/or summer salary will contribute to the
% stated goal(s) of your proposed project. You are eligible to apply for up to
% $50,000/year.

\input{preamble.tex}
% even page, right position (inner)
\rehead[]{Dav Clark -- \textbf{Budget} -- Data Science for Social Impact}
% odd page, left position (inner)
\lohead[]{Dav Clark -- \textbf{Budget} -- Data Science for Social Impact}

\begin{document}

After a discussion with Cathryn Carson, I'm asking for a \textbf{50/50 split for
    the duration of my current appointment in the D-Lab}, with the expectation
that I would either continue in the D-Lab or find additional support from
another organization on campus. My options for support are described more fully
in the \emph{funding} document. My current yearly salary is \$73,000 with a
benefit rate of 36.8\%. Half of that total is \$49,932. A 3\% increase
in the following year would be \$50,081.

However, as Cathryn notes in her thoughts on this issue (below, included at her
request), \emph{if I weren't constrained by the circumstances around the D-Lab
director transition}, I would propose one year of 100\% D-Lab (up front) followed
by 1 year of 100\% BIDS.

\vspace{2ex}
\noindent From Cathryn:

\begin{quote}
This is a hard one. I have a conflict of interest. On D-Lab's behalf I would
love to get 50\% of your salary back near-term. I'd use that money for doing more
great stuff in D-Lab. On the other hand, D-Lab probably also has an interest in
keeping you for 2 years -- but it would go against my stated approach to commit
the future director to that. On BIDS's behalf I would love to see you there for
2 years. Or, more pointedly, longer, since we're supposed to be building career
paths long-term, and you're one of the people who can make data science sing.

And that's leaving aside any considerations I might want to add in for your own
career satisfaction, sense of confidence in the university, or feeling of being
valued and rewarded for transformative work.

What I feel I can commit to for D-Lab is 50/50 split for the rest of your
current appointment in D-Lab. And the way the  power structure is set up in the
BIDS call for proposals, that's what decides it.

How about you 1) Put in a budget for 50/50 split for the rest of your current
appointment in D-Lab 2) Explain in the narrative that if you weren't constrained
by the circumstances around the D-Lab director transition, you would propose 1
year of 100\% D-Lab (up front) followed by 1 year of 100\% BIDS.  3) Copy this
message into the text of the narrative.
\end{quote}

\end{document}
