% How will the resources provided by BIDS be leveraged for greater impact in your research group?

\input{preamble.tex}
% even page, right position (inner)
\rehead[]{Dav Clark -- \textbf{Impact} -- Data Science for Social Impact}
% odd page, left position (inner)
\lohead[]{Dav Clark -- \textbf{Impact} -- Data Science for Social Impact}

\begin{document}

At the D-Lab, I interface with groups ranging from the Berkeley Institute for
Transparency in the Social Sciences (BITSS) to traditional social science
departments to UCSF/UCB purchasing. I'm “the data scientist” at the D-Lab. This
is an even more unwieldy proposition than being “the statistician” for the
social sciences (and humanities, and the purchasing department...). To meet this
challenge, I've networked with an enormous variety of organizations on campus.
These contacts can range from working to solve individual's immediate data
analysis problems to thinking about how to build campus-wide solutions that
really scale. Here are examples of things I'm already doing:

\begin{enumerate}
    \item Organizing the first publicly visible event for BIDS: the sold-out
        BIDS Software Carpentry Bootcamp.
        (\url{http://swcarpentry.github.io/2014-03-17-ucb}, Running March 17-18.)
    \item Coordinating with Lester Mackey (professor of statistics at Stanford)
        on creating a Bay Area hub for socially engaged data science:
        \url{statsforchange.org}.
    \item Further coordinating this initiative with the broader Data Science for
        Social Good program at the Urban Data Center at U Chicago:
        \url{dssg.io}. This program is led by Matt Gee and Rayid Ghani, who were
        instrumental in setting the theme for this years KDD conference, “Data
        Mining for Social Good.”
    \item Building a standardized computing environment that can provide
        consistent learning and research contexts anywhere from your laptop to
        “the cloud.” Here, I'm collecting input from staff and faculty in EECS,
        Statistics, Shared Research Computing, the D-Lab, and the iSchool in
        addition to advice from outside members of the data science community.
    \item Together with some amazing support from the community, I've revived
        and now organize the scientific Python meetings on campus. (Fernando
        used to run this before he got too busy with larger things!)
    \item Working with a variety of stakeholders on the development of
        high-quality curricula. I facilitate the D-Lab methods reading group
        (\url{dlab-berkeley.github.io/dlab-methods}), and have been working with
        educators inside and outside the D-Lab on planning curricula for
        traditionally under-served populations. Notably, I assisted Chris
        Holdgraf (graduate student in neuroscience) in writing his proposal to
        share a practical curriculum for statistical and computational
        training he's helped develop.
    \item Providing infrastructure to make John Canny's GPU accelerated text
        parser available to social scientists.
    \item Developing no-cost solutions to organize lightweight groups on campus
        using GitHub.
    \item Providing internally- and externally-facing training both in the
        D-Lab, and also for sister organizations on campus, like the computing
        topics session at the BITSS/ICPSR summer institute:
        \url{bitss.org/2014/02/13/announcing-bitss-summer-institute}.
\end{enumerate}

As a BIDS fellow, I will connect the BIDS community  with the groups I'm already
engaged with. If my experience is any judge, this will result in immense value
-- both in surfacing the really interesting, challenging problems, and sharing
effective, scalable solutions.

\end{document}
